\documentclass[11pt, letterpaper]{report}
\usepackage{amsmath,fullpage,graphicx}

\setlength{\parindent}{0em}

\newcommand{\nott}{{\sim}}

\begin{document}

{\textbf{Discrete Structures, Fall 2023, Homework 1}}

\vspace*{.1in}

You must write the solutions to these problems legibly on your own paper, with
the problems in sequential order, and with all sheets stapled together.

\vspace*{.1in}

\begin{enumerate}

	\item Convert the following sentences to logical statements using symbols assuming that ``$p$'', ``$b$''
	and ``$m$'' represent the propositions below.  
	
	$p$ = ``Morgan is taking a physics class.'' \\
	$b$ = ``Morgan is taking a biology class.'' \\
	$m$ = ``Morgan is taking a math class.''
	
	\begin{enumerate}
		\item Morgan is taking a biology class and a math class, but not a physics class.
		\item Morgan is taking a physics class, and either a biology or math class (but not both bio and math).
		\item Morgan is taking a physics class, and either a biology or math class (perhaps both bio and math).
		\item If Morgan is taking a biology class, then they are also taking a math class, but if Morgan is not taking a 
		biology class, then they are taking a physics class.
	\end{enumerate}
	
	\item For each of the sentences below, determine if the sentence is a statement.  If
	the sentence is a statement, tell whether it is true or false.
	\begin{enumerate}
		\item If a tree falls in the forest and no one is around to hear it, does it make a sound?
		\item If $2+2=5$, then I am the very model of a modern major-general.
		\item This statement refers to itself.
		\item This statement is false.
	\end{enumerate}
	
	\item Express the \textbf{negations} of the following statements in normal English sentences.
	\begin{enumerate}
		\item Sally is a computer science major and Sally's brother is a math major.
		\item Either the professor is late or my watch is fast.
	\end{enumerate}
	
	\item For each of the following statements, give the contrapositive, converse, and
	inverse statements (label them) in normal English, using the phrasing ``If \ldots, then \ldots''   
	You may change verb tenses to improve
	the grammar.  \textbf{Do not} use ``if'' or ``only if'' in the middle of your sentences.
	\begin{enumerate}
		\item ``If you conquer yourself, then you conquer the world.''\footnote{From \emph{Aleph}, by Paulo Coelho.}				
		\item I will be able to retire if I save enough money.
		\item You can go to the party only if you get good grades.
		
	\end{enumerate}
	
	\item Let $x$, $y$ and $z$ be statements.  Construct a complete truth table (meaning don't leave out any columns) for the 
	statement $(x \to \nott y) \land (\nott x \lor z)$.
	
	\item Are the statements $\nott (p \lor q)$ and $\nott p \lor \nott q$ logically equivalent?  Use a complete truth table to justify
	your answer, and explain (in English) why the truth table supports your answer.

\end{enumerate}


\end{document}

