\documentclass[12pt, letterpaper]{report}
\usepackage{amsmath,fullpage,graphicx,amssymb}

\setlength{\parindent}{0em}

\newcommand{\nott}{{\sim}}

\begin{document}

{\textbf{Discrete Structures, Fall 2023, Homework 3}}

\medbreak

You must write the solutions to these problems legibly on your own paper, with
the problems in sequential order, and with all sheets stapled together.

\begin{enumerate}

\item Complete the following proofs using the framework discussed in class.  Each line
        of your proof must be justified with a rule of inference or logical equivalence and appropriate line numbers.

\begin{enumerate}

\item
        \begin{tabular}[t]{ll}
        P1 & $s \land e$ \\
        P2 & $e \to b$ \\
        P3 & $(b \land \nott m) \to \nott s$ \\ \hline
        Prove: & $m$
        \end{tabular}           
        
                \item
                        \begin{tabular}[t]{ll}
                        P1 & $(p \to q) \land (r \to s)$ \\
                        P2 & $y$ \\
                        P3 & $(s \land q) \to \nott y$ \\ \hline
                        Prove: & $\nott p \lor \nott r$
                \end{tabular}
                
                                
                
                \item
                        \begin{tabular}[t]{ll}
                        P1 & $p \to q$ \\
                        P2 & $\nott q \lor r$ \\
                        P3 & $s \lor (y \land \nott r)$ \\ \hline
                        Prove: & $\nott s \to \nott (p \lor \nott y)$
                \end{tabular}
                
                
                
                \item
                        \begin{tabular}[t]{ll}
                        P1 & $a \land \nott d$ \\
                        P2 & $b \to (e \to d)$ \\ \hline
                        Prove: & $(a \to b) \to \nott e$
                \end{tabular}




\end{enumerate}

        \item Translate each of the following English sentences into formal language -- that is, using the symbols $\forall, \exists, \in,$ etc.  Use the following predicates: 
        
        $B(s)$ means ``$s$ is an business major,'' \\
        $C(s)$ means ``$s$ is a computer science
        major,'' and \\
        $M(s)$ means ``$s$ is a math major.'' 
        
         Use the domain $S = $ the set of
        all students at Rhodes College.
        
        \begin{enumerate}
                \item There is an business major who is also a math major.
                \item Every computer science major is also an business major.
                \item No computer science majors also major in business.
                \item Some people majoring in CS are also majoring in math.
                \item Some computer science majors are business majors as well, but some are not.  (Think carefully; this is tricky.)
        \end{enumerate}
        


\end{enumerate}

\end{document}
