\documentclass[12pt, letterpaper]{report}
\usepackage{amsmath,fullpage,graphicx,amssymb}

\setlength{\parindent}{0em}

\newcommand{\nott}{{\sim}}
\newcommand{\Z}{\mathbb{Z}}

\begin{document}

{\textbf{Discrete Structures, Fall 2017, Homework 5}}

\medbreak

\textit{You must write the solutions to these problems legibly on your own paper, with
the problems in sequential order, and with all sheets stapled together.}

\bigskip

For each statement below, state whether it is true or false. Then prove the statement if it is true,
or its negation if it is false.\medskip

Remember, an example may only be used to prove that an existential statement is true or a universal
statement is false. Any example or counter-example must include specific values for the variables
and enough algebra and justification to illustrate that the example proves what you are claiming
it proves.\medskip

You do not need to translate each statement into symbols first, though it is often useful to do so.

\begin{enumerate}

        \item The product of any two odd integers is odd.  % true
        
        \item For any integers $a$ and $b$, if $a-b$ is even, then $a$ and $b$ are both even.
        
        \item If $n$ is an even integer, then $n^2-2$ is even.  % true
        
        \item The sum of an integer and a rational number is rational.  % true
        
       	\item If $n$ and $m$ are rational numbers, then $n/m$ is a rational number.
                

\end{enumerate}
Hints: Remember, not all of these are necessarily true!  Use the step-by-step example proofs from the handouts last week
as a guide for these. 

\end{document}
